---
title: "linear_models"
author: "Jingyi Yao"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

set.seed(1)


```

## Model Fitting

  * interactions between variables can be specified using `*`.
  * intercept-only model `outcome ~ 1`
  * a model with no intercept `outcome ~ 0 + preidctors`
  * model using all available predictors `outcome ~ .`
  
```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(stars = review_scores_location / 2) %>% 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) %>% 
  filter(borough != "Staten Island") %>% 
  select(price, stars, borough, neighborhood, room_type)
```

  
R will treat categorical (factor) covariates appropriately and predictably: indicator variables are created for each non-reference category and included in your model, and the factor level is treated as the reference.

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
fit
```


changing reference categories won’t change “fit” or statistical sigificance, but can affect ease of interpretation.

```{r}
nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(
    borough = fct_infreq(borough),      # reorder factor levels by frequency
    room_type = fct_infreq(room_type))


fit = lm(price ~ stars + borough, data = nyc_airbnb)

fit

```


```{r}


```


```{r}


```


```{r}


```


```{r}


```


```{r}


```


```{r}


```


```{r}


```


```{r}


```


```{r}


```


```{r}


```


```{r}


```